A Safe and High-Performance Aliasing-Tolerant Parallel Programming Model

### 2012-11-14 doc/types-pldi2013/langdef.tex
hopefully last wave of edits, spell check

Figure~\ref{fig:langdef} defines Core Legion, a subset of the full Legion language
that illustrates the important issues.  
%Types include booleans, integers, tuples, and pointers.  
Pointers are annotated with a list of logical regions---a non-null pointer must point to a 
location that is contained in at least one of the regions. There is a special
{\em coloring} type, used to specify partitions of regions into subregions.
%Functions in Legion are named, accept one or more arguments and
%return a value of some type.  
Functions declare the region
access privileges and coherence they require.  Some function instances are {\em tasks}; a function may be executed
within the current task or run in parallel as a separate (sub)task---this decision is made
by the Legion task scheduler.  Function types are
universally quantified over all region names appearing in the type.
The other important type is a {\em region relationship}, an instance of which captures a value and
one or more regions satisfying associated constraints.  A region relation instance
can be written into the heap and later read and
unpacked, giving new local names to the regions contained in
the instance.

@@

Instead we use a bottom-up approach with big-step semantics.
Each expression has a clobber set $C$ of memory locations that might 
be modified by concurrent evaluations.  While a read from a location not in $C$ is {\em coherent}
(returning the most recently written value), the value read from a location $l$ in $C$ may be any
value.  Determining which value is deferred until all potentially-concurrent
modifications to $l$ are observable in the evaluation of a parent expression whose clobber set does
not include $l$ and a consistent interleaving of the subexpressions' memory traces is determined.
%With this formulation, the criteria for when a hierarchical scheduler may execute subexpressions in parallel can be
%made clear, but they are no longer constructive.  Instead, we will have to show that the scheduling algorithm
%used by the Legion runtime satisfies these constraints, guaranteeing the soundness of the Legion programming model.

@@

% FIXME The explanation of newcolor is confusing.
%
%The general form of a coloring being an abitrary mapping of locations to ``colors'' (integers) is shown in the
%operational semantics rules for the {\em newcolor} and {\em color} expressions.  
The {\em color} expression updates an existing coloring with a new color for a location, replacing any
existing coloring for that location.  The behavior of {\em newcolor} is subtler.
Legion supports first partitioning regions and then allocating data in the subregions.
%the {\em partition-then-allocate} model (see Section~\ref{sec:example}) in which the {\em new} expression may be called
%on subregions, 
Thus, each subregion resulting from a partitioning may need an arbitrary number of locations that are
unknown at partitioning time.  The semantics of newcolor allows this while
guaranteeing all explicit colorings performed by the program are honored.

@@

Finally, calls to subtasks evaluate actual arguments and create the subtasks's
environment (with potentially different names for logical regions).  A key point in Legion task calls is the
handling of relaxed coherence.  Any location that falls within
a logical region whose coherence has been relaxed (via {\em atomic} or {\em simult}) is added to the clobber set
$C'$, allowing non-deterministic evaluation within the subtask with respect to those locations.  All actual accesses
to those locations within the resulting memory trace ($E_{n+1}$) are marked with the corresponding coherence mode,
allowing the {\em valid\_interleave} check within the caller's environment to verify the consistency of the overall
memory trace.

@@

Our first result shows a well-typed expression accesses the heap in ways consistent with its static privileges.
A judgment $\privconsist{E}{\Phi}$ holds if memory operations in memory trace $E$ have types and
locations covered by privileges $\Phi$:
\[ 
\begin{array}{l}
\privconsist{E}{\Phi} \Leftrightarrow \forall \epsilon \in E. \\
\quad\quad (\epsilon = read(l, c, v) \Rightarrow \exists r. l \in M(r) \wedge \text{reads}(r) \in \Phi)\ \wedge \\
\quad\quad (\epsilon = write(l, c, v) \Rightarrow \exists r. l \in M(r) \wedge \text{writes}(r) \in \Phi)\ \wedge \\
\quad\quad (\epsilon = reduce_{id}(l, c, v) \Rightarrow \exists r. l \in M(r) \wedge \text{reduces}_{id}(r) \in \Phi)
\end{array}
\]

\noindent
%We must show $\privconsist{E}{\Phi}$ holds for any execution of an expression $e$ with static privileges $\Phi$.
As usual, the soundness claim is proven assuming the initial type and execution environments are consistent.
For our results, three consistency properties are needed:
\begin{itemize}
\item mapping consistency, written $\mapconsist{\Omega}$, guarantees a region mapping $M$ satisfies the region constraints $\Omega$% under which an expression was typed
\item local value consistency, written $\localconsist{L}{\Gamma}$, guarantees local values in $L$ have types consistent with the environment $\Gamma$
%under which the expression was typed
\item store consistency, written $\storeconsist{S}{H}$, guarantees locations in $S$ have values consistent with heap typing $H$
\end{itemize}

@@

An important insight from the proof of Theorem~\ref{thm:effects} is that 
it does not rely on the full definition of {\em valid\_interleave}.  Weakening 
{\em valid\_interleave} to {\em any\_interleave} (Figure~\ref{fig:validinterleave}), 
which  allows arbitrary interleavings of memory traces from subexpressions, 
is sufficient for the proof of soundness.  
The stronger constraints in {\em valid\_interleave}
(Figure~\ref{fig:validinterleave}) address the coherence of heap accesses;
{\em valid\_interleave} specifies permitted interleavings of memory operations for
the particular coherence modes on logical regions.

To determine whether an interleaving of two or more memory traces is valid, we consider three sets of addresses:
\begin{itemize}
\item exclusive locations ($l \in L_{excl}$) are those which have at least one access in exclusive mode in the traces
and are not in the clobber set.  For these locations, we require sequential execution semantics - all
reads to these locations see the effect of previous writes and reductions, and the resulting state of the store is
as if all writes and reductions were applied from each trace in order.
\item atomic locations ($l \in L_{atomic}$) are those which have at least one access in atomic mode in the traces and
are in neither $L_{excl}$ nor the clobber set.  For these locations, we allow permutations of the original
subexpression trace order.
\item for locations with only access in {\em simult} mode or in the clobber set, no constraints are
enforced.  The validity of the interleaving of these accesses is performed by some enclosing expression.
\end{itemize}

@@

To determine when parallel execution is safe, we start from the sequential execution trace and allow the reordering of 
adjacent heap operations that do not change the behavior of the application.  If we can show that it is safe to 
reorder any pair of operations that come from two different constituent traces, then any interleaving of the constituent
traces will be equivalent to a sequential execution and parallel execution is safe.  To
efficiently discover these cases at runtime, we require a test that can determine this property prior
to the actual execution of the tasks that create the traces.  We show that a test based on the subtask privileges and 
the current region mapping can soundly predict when this property will hold.
%If we can show that reordering is allowed
%between any two operations that come from different constituent traces, any interleaving of the constituent traces will
%be equivalent to the serialized execution and parallel execution was safe.  Finally, to allow the safety of parallel
%execution to be determined {\em before} the execution is actually performed, we show that a test based on the subtask
%privileges and the current region mapping can soundly predict when this property will hold.
We begin by defining a {\em non-interference} operator on two memory operations
$\epsilon_1 = op_1(l_1, c_1, v_1)$ and $\epsilon_2 = op_2(l_2, c_2, v_2)$:
$$
\begin{array}{l@{ }l}
\epsilon_1\ \nonint\ \epsilon_2 \Leftrightarrow & (op_1 = read \wedge op_2 = read)\ \vee \\
& (op_1 = reduce_{id_1} \wedge op_2 = reduce_{id_2} \wedge id_1 = id_2) \vee \\
& l_1 \not= l_2
\end{array}
$$
Reads have no side effects, and cannot change what another read returns.  The safety of the second case
follows from the requirement that reduction operations be commutative.  
Finally, accesses to different locations cannot affect each other.
We next show that this non-interference operator preserves the validity of an interleaving
of two or more traces when reordering a pair of memory operations.
\begin{lem}
\label{lem:nonintswap}
\rm
Let $S$ be a store, $C$ a clobber set, $\oton{E}, E'_a, E'_b$ memory traces, and
$\epsilon_1, \epsilon_2$ be two memory operations from $E_i$ and $E_j$ ($i \not= j$).  Then,
%%$$\begin{array}{l}
\begin{align*}
valid&\_interleave(S, C, E'_a \dplus [ \epsilon_1, \epsilon_2 ] \dplus E'_b, \oton{E}) \wedge \epsilon_1 \nonint \epsilon_2 \\
&\Rightarrow valid\_interleave(S, C, E'_a \dplus [ \epsilon_2, \epsilon_1 ] \dplus E'_b, \oton{E}).
\end{align*}
%%\end{array}$$
\end{lem}
%The proof follows from the definitions of {\em coherent} and {\em seq\_equiv} in Figure~\ref{fig:validinterleave}.

@@

$
\begin{array}{l@{}l}
\multicolumn{2}{l}{priv_1(r_1) \nonint[M] priv_2(r_2) \Leftrightarrow} \\
& (priv_1 = \text{reads} \wedge priv_2 = \text{reads})\ \vee \\
& (priv_1 = \text{reduces}_{id_1} \wedge priv_2 = \text{reduces}_{id_2} \wedge id_1 = id_2)\ \vee \\
& (r_1 * r_2)\ \vee \\
& (M(r_1) \cap M(r_2) = \emptyset) \\
\\
\multicolumn{2}{l}{\Phi_1 \nonint[M] \Phi_2 \Leftrightarrow \displaystyle\bigwedge_{\phi_1 \in \Phi_1, \phi_2 \in \Phi_2} \phi_1 \nonint[M] \phi_2}
\end{array}$

@@

\texcomment{
Let $e_1$ and $e_2$ be well-typed Legion expressions using privileges $\Phi_1$ and
$\Phi_2$.  Now let $e_1'$ be some subexpression of $e_1$ and $e_2'$ be some 
subexpression of $e_2$, with corresponding privileges $\Phi_1'$ and $\Phi_2'$.
By Theorem~\ref{thm:parallelexec}, if the runtime can verify $\Phi_1 \nonint[M] \Phi_2$, then 
expressions $e_1$ and $e_2$ may be evaluated in parallel.  The evaluation of the subexpressions
$e_1'$ and $e_2'$ will therefore be allowed to run in parallel, and we must show that this is
permissible.  We do so by proving that any memory traces that can result from the evaluation
of the subexpression must be non-interfering (allowing parallel execution by
Lemma~\ref{lem:noninteffects}).
}
\begin{thm}
\label{thm:hiersched}
\rm
Let $e_1$ and $e_2$ be well-typed expressions using privileges $\Phi_1$ and
$\Phi_2$ respectively, where $\Phi_1 \nonint[M] \Phi_2$.  Let $e_1'$ be a subexpression of $e_1$ and $e_2'$ be a 
subexpression of $e_2$.  Any memory traces $E_1'$ of $e_1'$ and $E_2'$ of $e_2'$ resulting from evaluation of
$e_1$ and $e_2$ (with the usual consistent $M$, $L$, $H$, and $S$) are 
non-interfering.
\end{thm}
\texcomment{
\begin{proof}
Let $E_1$ be the result of an evaluation of $e_1$, with $E_1'$ being the memory trace from the
evaluation of $e_1'$ within the expression tree.  $E_1$ will be formed from a tree of valid
interleavings, each of which must include all the memory operations of each constituent trace.
By induction over the number of intermediate subexpressions between $e_1$ and $e_1'$, we
can show that any memory operation $\epsilon_1'$ that is in $E_1'$ must also be in $E_1$.
Similarly, any memory operation $\epsilon_2'$ that is in $E_2'$ must also be in $E_2$.  We have
$E_1 \nonint E_2$ from Lemma~\ref{lem:nonintpriv}, so and therefore $\epsilon_1' \nonint \epsilon_2'$.  Since
this holds for all $\epsilon_1'$ in $E_1'$ and all $\epsilon_2'$ in $E_2'$, we have $E_1' \nonint E_2'$.
\end{proof}
} The Legion task scheduler uses Theorem~\ref{thm:hiersched} as follows: 
sibling function calls (those invoked within the same
function body) $e_1$ and $e_2$ are checked for non-interference of their (dynamic)
privileges.  Since $e_1$ and $e_2$ are guaranteed to be created on the node on which their parent
is running, no communication is required to perform the non-interference
test.  If they interfere they are executed in program order or serialized
depending on their coherence specifications; otherwise they are
considered for execution as parallel subtasks.    If $e_1$ and $e_2$ are 
determined to be non-interfering and are scheduled in parallel on different 
remote processors then Theorem~\ref{thm:hiersched}
guarantees that there is no communication required between $e_1$ and $e_2$
to perform non-interference tests between their sub-tasks.  Therefore,
the runtime requires no communication for scheduling.

>>>>

Figure~\ref{fig:langdef} defines Core Legion, a subset of the full Legion language
that illustrates the important issues.  
%Types include booleans, integers, tuples, and pointers.  
Pointer types are annotated with a list of logical regions---a non-null pointer value must point to a 
location that is contained in at least one of the regions. There is a special
{\em coloring} type, used to specify partitions of regions into subregions.
%Functions in Legion are named, accept one or more arguments and
%return a value of some type.  
Functions declare the region
access privileges and coherence they require.  Some function instances are {\em tasks}; a function may be executed
within the current task or run in parallel as a separate (sub)task---this decision is made
by the Legion task scheduler.  Function types are
universally quantified over all region names appearing in the type.
The other important type is a {\em region relationship}, an instance of which captures a value and
one or more regions satisfying associated constraints.  A region relation instance
can be written into the heap and later read and
unpacked, giving new local names to the regions contained in
the instance.

@@

Instead we use a bottom-up approach with big-step semantics.
Each expression has a clobber set $C$ of memory locations that might 
be modified by concurrent evaluations.  Reads of locations in $C$ may be
chosen to return a value other than what was most recently written by the current
expression, as long a
valid interleaving of the operations with other concurrent operations can be found within
an enclosing expression.
%% While a read from a location not in $C$ is {\em coherent}
%% (returning the most recently written value), the value read from a location $l$ in $C$ may be any
%% value.  Determining which value is deferred until all potentially-concurrent
%% modifications to $l$ are observable in the evaluation of a parent expression whose clobber set does
%% not include $l$ and a consistent interleaving of the subexpressions' memory traces is determined.
%With this formulation, the criteria for when a hierarchical scheduler may execute subexpressions in parallel can be
%made clear, but they are no longer constructive.  Instead, we will have to show that the scheduling algorithm
%used by the Legion runtime satisfies these constraints, guaranteeing the soundness of the Legion programming model.

@@

% FIXME The explanation of newcolor is confusing.
%
%The general form of a coloring being an abitrary mapping of locations to ``colors'' (integers) is shown in the
%operational semantics rules for the {\em newcolor} and {\em color} expressions.  
The {\em color} expression updates an existing coloring with a new color for a location, replacing any
existing coloring for that location.  The behavior of {\em newcolor} is subtler.
Legion supports first partitioning regions and then allocating data in the subregions.
%the {\em partition-then-allocate} model (see Section~\ref{sec:example}) in which the {\em new} expression may be called
%on subregions, 
Thus, each subregion resulting from a partitioning may need an arbitrary number of locations that are
unknown at partitioning time.  The semantics of newcolor allows extra locations to be placed in every
subregion while
guaranteeing all explicit colorings performed by the program are honored.

@@

Finally, calls to subtasks evaluate actual arguments and create the subtasks's
environment (with potentially different names for logical regions).  A key point in Legion task calls is the
handling of relaxed coherence.  Any location that falls within
a logical region whose coherence has been relaxed (via {\em atomic} or {\em simult}) is added to the
subtask's clobber set $C'$, allowing non-deterministic evaluation within the subtask with respect to
those locations.  All actual accesses to those locations within the resulting memory trace ($E_{n+1}$)
are marked with the corresponding coherence mode, allowing the {\em valid\_interleave} check within
the caller's environment to verify the consistency of the overall memory trace.

@@

Our first result shows a well-typed expression accesses the heap in ways consistent with its static privileges.
A judgment $\privconsist{E}{\Phi}$ holds if memory operations in memory trace $E$ have types and
locations covered by privileges $\Phi$:
\[ 
\begin{array}{l}
\privconsist{E}{\Phi} \Leftrightarrow \forall \epsilon \in E. \\
\quad\quad (\epsilon = read(l, c, v) \Rightarrow \exists r, l \in M(r) \wedge \text{reads}(r) \in \Phi)\ \wedge \\
\quad\quad (\epsilon = write(l, c, v) \Rightarrow \exists r, l \in M(r) \wedge \text{writes}(r) \in \Phi)\ \wedge \\
\quad\quad (\epsilon = reduce_{id}(l, c, v) \Rightarrow \exists r, l \in M(r) \wedge \text{reduces}_{id}(r) \in \Phi)
\end{array}
\]

\noindent
%We must show $\privconsist{E}{\Phi}$ holds for any execution of an expression $e$ with static privileges $\Phi$.
As usual, the soundness claim is proven assuming the initial type and execution environments are consistent.
For our results, three consistency properties are needed:
\begin{itemize}
\item mapping consistency, written $\mapconsist{\Omega}$, guarantees a region mapping $M$ satisfies the region constraints $\Omega$% under which an expression was typed
\item local value consistency, written $\localconsist{L}{\Gamma}$, guarantees local values in $L$ have types consistent with the environment $\Gamma$ (with logical regions mapped to physical regions via $M$)
%under which the expression was typed
\item store consistency, written $\storeconsist{S}{H}$, guarantees locations in $S$ have values consistent with heap typing $H$
\end{itemize}

@@

An important insight from the proof of Theorem~\ref{thm:effects} is that 
it does not rely on the full definition of {\em valid\_interleave}.  Weakening 
{\em valid\_interleave} to {\em any\_interleave} (Figure~\ref{fig:validinterleave}), 
which  allows arbitrary interleavings of memory traces from subexpressions, 
is sufficient for the proof of soundness.  
The stronger constraints address the coherence of heap accesses;
{\em valid\_interleave} specifies permitted interleavings of memory operations for
the particular coherence modes on logical regions.

To determine whether an interleaving of two or more memory traces is valid, we consider three sets of addresses:
\begin{itemize}
\item exclusive locations ($l \in L_{excl}$) are those which have at least one access in exclusive mode in the traces
and are not in the clobber set.  For these locations, we require sequential execution semantics - all
reads to these locations see the effect of previous writes and reductions, and the resulting state of the store is
as if all writes and reductions were applied from each trace in order.
\item atomic locations ($l \in L_{atomic}$) are those which have at least one access in atomic mode in the traces and
are in neither $L_{excl}$ nor the clobber set.  For these locations, we allow permutations of the original
subexpression trace order.
\item for locations with only access in {\em simult} mode or in the clobber set, no constraints are
enforced.  The validity of the interleaving of these accesses is determined within the context of some enclosing expression.
\end{itemize}

@@

To determine when parallel execution is safe, we start from the sequential execution trace and allow the reordering of 
adjacent heap operations that do not change the behavior of the application.  If we can show that it is safe to 
reorder any pair of operations that come from two different constituent traces, then any interleaving of the constituent
traces will be equivalent to a sequential execution and parallel execution is safe.  To
efficiently discover these cases at runtime, we require a test that can determine this property prior
to the actual execution of the tasks that create the traces.  We show that a test based on the subtask privileges and 
the current region mapping can soundly predict when this property will hold.
%If we can show that reordering is allowed
%between any two operations that come from different constituent traces, any interleaving of the constituent traces will
%be equivalent to the serialized execution and parallel execution was safe.  Finally, to allow the safety of parallel
%execution to be determined {\em before} the execution is actually performed, we show that a test based on the subtask
%privileges and the current region mapping can soundly predict when this property will hold.
We begin by defining a {\em non-interference} operator on two memory operations
$\epsilon_1 = op_1(l_1, c_1, v_1)$ and $\epsilon_2 = op_2(l_2, c_2, v_2)$:
$$
\begin{array}{l@{ }l}
\epsilon_1\ \nonint\ \epsilon_2 \Leftrightarrow & (op_1 = read \wedge op_2 = read)\ \vee \\
& (op_1 = reduce_{id_1} \wedge op_2 = reduce_{id_2} \wedge id_1 = id_2) \vee \\
& l_1 \not= l_2
\end{array}
$$
Reads have no side effects, and cannot change what another read returns.  The safety of the second case
follows from the requirement that reduction operations be commutative.  
Finally, accesses to different locations cannot affect each other.
Therefore, an adjacent pair of non-interfering memory operations in a memory trace can be reordered while
preserving the validity of an interleaving.
%We next show that this non-interference operator preserves the validity of an interleaving
%of two or more traces when reordering a pair of memory operations.
\begin{lem}
\label{lem:nonintswap}
\rm
Let $S$ be a store, $C$ a clobber set, $\oton{E}, E'_a, E'_b$ memory traces, and
$\epsilon_1, \epsilon_2$ be two memory operations from $E_i$ and $E_j$ ($i \not= j$).  Then,
%%$$\begin{array}{l}
\begin{align*}
valid&\_interleave(S, C, E'_a \dplus [ \epsilon_1, \epsilon_2 ] \dplus E'_b, \oton{E}) \wedge \epsilon_1 \nonint \epsilon_2 \\
&\Rightarrow valid\_interleave(S, C, E'_a \dplus [ \epsilon_2, \epsilon_1 ] \dplus E'_b, \oton{E}).
\end{align*}
%%\end{array}$$
\end{lem}
%The proof follows from the definitions of {\em coherent} and {\em seq\_equiv} in Figure~\ref{fig:validinterleave}.

@@

$\begin{array}{@{}l@{}l@{}}
\multicolumn{2}{@{}l}{priv_1(r_1) \nonint[M] priv_2(r_2) \Leftrightarrow} \\
\hspace*{0.5cm} & (priv_1 = \text{reads} \wedge priv_2 = \text{reads})\ \vee \\
& (priv_1 = \text{reduces}_{id_1} \wedge priv_2 = \text{reduces}_{id_2} \wedge id_1 = id_2)\ \vee \\
& (r_1 * r_2)\ \vee \\
& (M(r_1) \cap M(r_2) = \emptyset) \\
\\
\multicolumn{2}{@{}l}{\Phi_1 \nonint[M] \Phi_2 \Leftrightarrow \displaystyle\bigwedge_{\phi_1 \in \Phi_1, \phi_2 \in \Phi_2} \phi_1 \nonint[M] \phi_2}
\end{array}$

@@

\texcomment{
Let $e_1$ and $e_2$ be well-typed Legion expressions using privileges $\Phi_1$ and
$\Phi_2$.  Now let $e_1'$ be some subexpression of $e_1$ and $e_2'$ be some 
subexpression of $e_2$, with corresponding privileges $\Phi_1'$ and $\Phi_2'$.
By Theorem~\ref{thm:parallelexec}, if the runtime can verify $\Phi_1 \nonint[M] \Phi_2$, then 
expressions $e_1$ and $e_2$ may be evaluated in parallel.  The evaluation of the subexpressions
$e_1'$ and $e_2'$ will therefore be allowed to run in parallel, and we must show that this is
permissible.  We do so by proving that any memory traces that can result from the evaluation
of the subexpression must be non-interfering (allowing parallel execution by
Lemma~\ref{lem:noninteffects}).
}
\begin{thm}
\label{thm:hiersched}
\rm
Let $e_1$ and $e_2$ be well-typed expressions using privileges $\Phi_1$ and
$\Phi_2$ respectively, where $\Phi_1 \nonint[M] \Phi_2$.  Let $e_1'$ be a subexpression of $e_1$ and $e_2'$ be a 
subexpression of $e_2$.  Any memory traces $E_1'$ of $e_1'$ and $E_2'$ of $e_2'$ resulting from evaluation of
$e_1$ and $e_2$ (with the usual consistent $M$, $L$, $H$, and $S$) are 
non-interfering.
\end{thm}
\texcomment{
\begin{proof}
Let $E_1$ be the result of an evaluation of $e_1$, with $E_1'$ being the memory trace from the
evaluation of $e_1'$ within the expression tree.  $E_1$ will be formed from a tree of valid
interleavings, each of which must include all the memory operations of each constituent trace.
By induction over the number of intermediate subexpressions between $e_1$ and $e_1'$, we
can show that any memory operation $\epsilon_1'$ that is in $E_1'$ must also be in $E_1$.
Similarly, any memory operation $\epsilon_2'$ that is in $E_2'$ must also be in $E_2$.  We have
$E_1 \nonint E_2$ from Lemma~\ref{lem:nonintpriv}, so and therefore $\epsilon_1' \nonint \epsilon_2'$.  Since
this holds for all $\epsilon_1'$ in $E_1'$ and all $\epsilon_2'$ in $E_2'$, we have $E_1' \nonint E_2'$.
\end{proof}
}

The Legion task scheduler uses Theorem~\ref{thm:hiersched} as follows: 
sibling function calls (those invoked within the same
function body) $e_1$ and $e_2$ are checked for non-interference of their (dynamic)
privileges.  Since $e_1$ and $e_2$ are called on the parent task's node, no communication
is required to perform the non-interference
test.  If they interfere they are executed in program order or serialized
depending on their coherence specifications; otherwise they are
considered for execution as parallel subtasks.    If $e_1$ and $e_2$ are 
determined to be non-interfering and are scheduled in parallel on different 
remote processors then Theorem~\ref{thm:hiersched}
guarantees that there is no communication required between $e_1$ and $e_2$
to perform non-interference tests between their sub-tasks.  Therefore,
the runtime requires no communication for scheduling.

### 2012-11-14 doc/types-pldi2013/langdef_full.tex
fixed typos found by Elliott - tried to tweak multicoloring and clobber set paragraphs
hopefully last wave of edits, spell check

\noindent Two additional properties are proven for each subexpression:
\begin{itemize}
\item $\valueconsist{v}{T}$, the expression yields a value of the right type
\item $\traceconsist{E}$, values written to the store match the heap types of their locations
\end{itemize}
Figure~\ref{fig:constprop} gives the precise definition of these
properties.\footnote{Note that in the definition of
$\mapconsist{\Omega}$, the quantified variables $r_i$ and $r_j$
range over both logical and physical regions.  Since $M$'s domain is
logical regions, we extend $M$ to constraints with mixed logical and
physical regions in the standard way by defining
$M \llbracket \rho \rrbracket = \rho$.  And although the definition of
$\traceconsist{E}$ refers to $M$ and $\Gamma$, the restriction that reduction
functions must be monomorphic and visible in all functions means that 
$\traceconsist{E}$ will be insensistive to the particular $M$ and $\Gamma$.}

>>>>

\noindent Two additional properties are proven for each subexpression:
\begin{itemize}
\item $\valueconsist{v}{T}$, the expression yields a value of the right type
\item $\traceconsist{E}$, values written to the store match the heap types of their locations
\end{itemize}
Figure~\ref{fig:constprop} gives the precise definition of these
properties.\footnote{Note that in the definition of
$\mapconsist{\Omega}$, the quantified variables $r_i$ and $r_j$
range over both logical and physical regions.  Since $M$'s domain is
logical regions, we extend $M$ to constraints with mixed logical and
physical regions in the standard way by defining
$M \llbracket \rho \rrbracket = \rho$.  And although the definition of
$\traceconsist{E}$ refers to $M$ and $\Gamma$, the restriction that reduction
functions must be monomorphic and visible in all functions means that 
$\traceconsist{E}$ will be insensitive to the particular $M$ and $\Gamma$.}
